### 2015-07-19 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Add reply to reviewer's comments

REPLY TO 1: We were trying to resolve the atomistic resolution structure of glycerol backbone 
and choline by using the available atomistic resolution models. In all of the used models the 
most recent version published by the developers was intended to be used in NPT ensemble with 
zero tension. The improvements of the models and testing of influence of changes in parameters, 
including the ensemble, is beyond the scope of the work. Further, it is not very likely that the 
change in the ensemble would significantly change the glycerol backbone or choline structure.

REPLY TO 2: It is known experimentally that the glycerol backbone order parameters are essentially 
the same for PC, PE and PG lipids, as written in the manuscript (page 15): 
"Since the glycerol backbone order parameters are very similar irrespective of the headgroup 
chemistry (PC,PE and PG) or lipid environment,26 it is reasonable to assume that the stereospecifity 
measured for the PE lipids holds also for the PC lipids." 
It is extremely unlikely scenario that the stereospecific labeling would be different while the 
order parameters would be numerically equal. Thus the assumption has quite strong experimental justification.

REPLY TO 3: The quantification used is explained in the figure caption. MAYBE MARKUS SHOULD CLARIFY THIS A LITTLE BIT? 
It is based on the distance between simulated and experimental order parameters. The figure is made 
and color coding is used to make the message in Fig. 2 more easily readable, thus all the details are 
also available for the readers who are willing to take a closer look. If the force fields are ranked based 
on the distance of the results from the experimental values, the order of the force fields in the figure will 
not change. The color coding may change if another researcher prefers some other coding. Since all the data 
is available, anyone is free to do that.

REPLY TO 4: As written in the manuscript there is actually also previous experimental support, in addition 
to the current results, for the dihedral distributions suggested in the manuscript. About g3-g2-g1-O dihedral 
it is written on page 19: 
"In conclusion, models preferring the anti conformation for this dihedral give more 
realistic order parameters and this is in agreement with previous crystal structure and 1H NMR studies.19–21,23–25" 
and about N-b-a-O dihedral it is writte on page 20: 
"This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies19–22 which suggest that this dihedral is in the gauche conformation in the absence of ions."

REPLY TO 4 (a): We do not know that the other models have exatcly the correct dihedral distribution 
and we do not claim this in the manuscript. Actually they cannot be all correct since they are not 
the exactly same and also the order parameters are not the same and none of them really satisfy the 
experiments, as discussed in the manuscript. However, as said in the manuscript and in the above reply, 
there is strong evidence that dihedrals in other models are significantly more realistic than in Berger.

The justification for usage of phrase "biologically relevant conditions" in the manuscript is given 
in the sentence on page 4-5: 
"Importantly, the structural parameters for the glycerol backbone are similar for various biologically 
relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
in various environments,26 and the structural parameters for the choline headgroup are similar in model 
membranes and real cells (mouse fibroblast L-M cell).27". 
Since the order parameters for these parts are the same in real cells (at least E. Coli and mouse 
fibroblast L-M cell) as in model systems, most likely the structure is the same as well.

REPLY TO 4 (b): Yes, they can. However, if molecular dynamics simulation model would reproduce all 
the experimental parameters correctly (including order parameters) it would be correct with very high 
probability, and would give a valid interpretation for experiments, as stated several times in the 
manuscript, for example on page 21: 
"If the improved atomistic molecular dynamics simulation model reproduced the order parameters 
and other experimental observables (like chemical shift anisotropy) with experimental accuracy, 
it would give an interpretation for the atomistic resolution structure of the glycerol backbone and choline.10–13,15,16,18".

REPLY TO 4 (c): As stated in the replies above, there is also previous experimental evidence for 
the conclusions made about dihedral distributions and these conclusions can be extended for the 
real cells as well. However, it is true that, in principle one cannot exclude 
other dihedrals than these two without further evidence. However, the previous experiments supports 
the conclusions and the main purpose of this analysis here is to  (from page 17-18) 
"demonstrate that 1) the differences in order parameters indicate significantly different 
structural sampling strongly correlating with the dihedral angles of the related bonds, 
and that 2) the comparison between experimental and simulated order parameters can be used 
to exclude nonrealistic structural samping in molecular dynamics simulations." 
And further from page 19 
"We note that performing a similar comparison through all the dihedrals in all the 13 
models would probably give highly useful information on how to improve the accuracy of the models 
yet this is beyond the scope of the current report."

REPLY TO 5: This is indeed relevant work and we were aware of it. The reference was accidentally 
left out from the manuscript but it WILL BE ADDED. We thank the reviewer for pointing this out.

REPLY TO 6: This is indeed relevant for the simulations with cholesterol since the glycerol tilt 
may also explain the minor changes in the order parameters when adding cholesterol. 
WE SHOULD CALCULATE THIS FROM CHOLESTEROL SIMULATIONS WITH CHARMM SINCE THOSE ARE USED FOR THE 
DIHEDRAL ANALYSIS IN SI. MAYBE HUBERT COULD DO THIS? 
As a function of hydration level unfortunately (from page 24): 
"The response of the glycerol backbone seems to be more subtle than that of the choline 
headgroup; none of the four models reproduced the experimental trends upon dehydration 
with enough accuracy to invite a structural interpretation."

REPLY TO 7: The scope of this manuscript has been limited on purpose on the glycerol backbone 
and choline segments to avoid exhausting the reader, thus this analysis is beyond the scope 
of the work. The acyl chain region will be discussed in the context of the NMRLipids III 
project (see http://nmrlipids.blogspot.fi/p/about.html).