### 2015-10-11 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt

Reviewer 2.

We thank for the very positive feedback. The errors in English are
now fixed as suggested.

Reviewer 1.

1.
Author reply:

2.
Author reply:

All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.

This size was agreed on based on the high experimental repeatibility of the order
parameters that we discuss in the text (p12): "Almost all the reported values are
within a variation of ±0.02 [...] for all fully hydrated PC bilayers, regardless
of variation in their acyl chain composition and temperature."

To make the (equal) size of the sweet spots clear to the reader, we now explicitly
point out this size when introcuding them in the text (p15):

"Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
subjective sweet spots (light blue areas **spanning 0.04 units**), within which
we expect the calculated values of the order parameters of a well-performing
force field to fall."

The added words are those between the stars (**).

3.
Author reply:
The title of our manuscript is:
"Towards atomistic resolution structure of
phosphatidylcholine headgroup and glycerol
backbone at different ambient conditions"
in the introduction we write:
"Phospholipid molecules are composed of hydrophobic acyl chains connected
by a glycerol backbone to a hydrophilic headgroup ..."
This is widely used naming convention for phospholipid molecules.
According to the title and the naming convetion used in the manuscript, 
the beginning of acyl chains are not the part of headgroup and are 
beoynd the scope defined in the title. Their presence at the interface
or their chemical environment do not change this.

We are not willing to extend the scope of the manuscript since it
would signifincantly increase already massive information content and 
length. Instead, the acyl chains order parameters will be discussed in 
the context of the NMRLipids III project, see 
http://nmrlipids.blogspot.fi/p/about.html.