### 2015-10-13 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Add reply to comment 1 by the reviewer

The original reviewer comments are not written in this publicly available
file we do not have permission to publish those. The original comments
will be added to the version which will be submitted to the JPCB.

Reviewer 2.

We thank for the very positive feedback. The errors in English are
now fixed as suggested.

Reviewer 1.

1.
Author reply:
In the publications mentioned by the reviewer it is shown that the
the membrane properties are affected by interfacial tension rather than the ensemble. 
This is also explicitly said in the abstract of
J. Chem. Phys. 111, 1281-1287
"Equivalence between the constant area and constant surface tension ensembles is
investigated by comparing the present simulations with earlier work from our laboratories and we
find simulation results to depend much more strongly on the specified surface area or surface tension
than on the ensemble employed"
It should be also noted that the total lengths of the simulations are around 1ns 
in the mentioned publications while our simulations are 1-2 orders of magnitude longer. 
Thus, even though the dependence on the ensemble would be observed in these short simulations
it could very well be due to insufficient sampling.

As we have already pointed out we have restricted our work to the available models and
the improvement of these models has been left for further studies. Since all the developers
of the tested models have used NPT ensemble in zero tension, there seems to be consensus on 
this.

It is also much more likely that the force field parameters affect the glycerol backbone and
choline structures much more than the used ensemble or interfacial tension. We have 
left the more detailed discussion on the force field parameters for further studies, 
thus the same should be done for the discussion about ensemble and surface tension which
is most likely less relevant.

2.
Author reply:
All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.

This size was agreed on based on the high experimental repeatibility of the order
parameters that we discuss in the text (p12): "Almost all the reported values are
within a variation of ±0.02 [...] for all fully hydrated PC bilayers, regardless
of variation in their acyl chain composition and temperature."

To make the definition of the sweet spots clear to the reader, we now explicitly
point this out when introducing them in the text (p15):

"Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
subjective sweet spots (light blue areas **spanning 0.04 units around the average
of the extremal experimental values**), within which
we expect the calculated values of the order parameters of a well-performing
force field to fall."

The added words are those between the stars (**).

3.
Author reply:
The title of our manuscript is:
"Towards atomistic resolution structure of
phosphatidylcholine headgroup and glycerol
backbone at different ambient conditions"
in the introduction we write:
"Phospholipid molecules are composed of hydrophobic acyl chains connected
by a glycerol backbone to a hydrophilic headgroup ..."
This is widely used naming convention for phospholipid molecules.
According to the title and the naming convetion used in the manuscript, 
the beginning of acyl chains are not the part of headgroup and are 
beoynd the scope defined in the title. Their presence at the interface
or their chemical environment do not change this.

We are not willing to extend the scope of the manuscript since it
would signifincantly increase already massive information content and 
length. Instead, the acyl chains order parameters will be discussed in 
the context of the NMRLipids III project, see 
http://nmrlipids.blogspot.fi/p/about.html.